---
title: ""
output: html_document
date: ""
---

```{r setup, include=FALSE}

library(ggplot2)
library(ggsci)
library(dplyr)

my_theme <- function()
{
  res <- theme(
    plot.title = element_text(hjust=0),
    plot.title.position = "plot",
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(linetype = "dotted", color = "grey"),
    panel.background = element_rect(fill = "white"),
    legend.position = "bottom")
  res
}

load("../p_origin.Rdata")

```

### Chr1 maternal folic acid interaction  

Almost all SNPs in this region had GTEx entries. GTEx data is shown for the highlighted SNPs. The RSIDs on the x-axis of the eQTL plots are in order of BP position (so they match left-to-right the highlighted SNPs)

```{r, echo=FALSE}
load("../mat.Rdata")
load("../m_LD1.Rdata")

mat <- mat[mat$chr == 1,]
chr1_GTEx <- read.table("chr1_GTEx.txt", header=TRUE, sep='\t', quote="")
positions <- sapply(strsplit(chr1_GTEx[,3], ":"), "[[", 2)
positions <- as.numeric(positions)
chr1_GTEx <- chr1_GTEx[positions %in% mat$position,]
mat$rsid[mat$position %in% positions] <- unique(chr1_GTEx$RS.ID)

m_LD1 <- m_LD1[rownames(m_LD1) %in% mat$snp, colnames(m_LD1) %in% mat$snp]
lead  <- mat %>% slice_min(folate) %>% pull(snp)
mat$LD<- m_LD1[,lead]^2

keep <- (-log10(mat$folate) > 2) & (mat$LD > 0.8)

p1 <- ggplot(mat, aes(position, -log10(folate))) +
    geom_point(color="grey") +
    geom_point(data=mat[keep,], aes(position, -log10(folate), color=LD)) +
    # geom_text(data=mat[keep,], aes(position, y=3, label=rsid), nudge_y=0.5, angle=30) +
    scale_colour_stepsn(breaks = c(0.4, 0.6, 0.8), colours = pal_locuszoom("default")(5)) + 
    ylim(c(0,5)) + labs(y="-log10(pval)") +my_theme() +
    theme(axis.text.x=element_blank())

chr1_GTEx <- chr1_GTEx[chr1_GTEx$RS.ID %in% mat[keep,]$rsid,]
chr1_GTEx$RS.ID <- factor(chr1_GTEx$RS.ID, levels=unique(chr1_GTEx$RS.ID))

p2 <- ggplot(chr1_GTEx, aes(x=RS.ID, y=-log10(P.value), fill=Tissue)) + 
    geom_col(position=position_dodge2()) + facet_wrap(vars(Gene.Symbol), ncol =2) +
    theme(axis.text.x =element_text(angle=90),
          axis.title.x=element_blank(),
          legend.position = "bottom") +
    my_theme()

p1
p2  + labs(title="GTEx expression data for highlighted SNPs")
# cowplot::plot_grid(p1,p2, ncol=1, rel_heights = c(1,3), labels=c("A","B"))
```

### Chr2 133Mb Parent-of-origin Smoking interaction  

Only 3 SNPs in this region were present in GTEx v8

```{r, echo=FALSE}
load("../p_origin.Rdata")
load("../p_LD1.Rdata")

chr2 <- p_origin[p_origin$chr== 2, ]
chr2_GTEx <- read.table("chr2_GTEx.txt", header=TRUE, sep='\t', quote="")
positions <- sapply(strsplit(chr2_GTEx[,3], ":"), "[[", 2)
positions <- as.numeric(positions)
chr2_GTEx <- chr2_GTEx[positions %in% chr2$position,]
chr2$rsid[chr2$position %in% positions] <- unique(chr2_GTEx$RS.ID)

keep <- chr2$rsid %in% chr2_GTEx$RS.ID

p1 <- ggplot(chr2, aes(position, -log10(smoke))) +
    geom_point(color="grey") +
    geom_point(data=chr2[keep,], aes(position, -log10(smoke)), color="red") +
    # geom_text(data=mat[keep,], aes(position, y=3, label=rsid), nudge_y=0.5, angle=30) +
    # scale_colour_stepsn(breaks = c(0.4, 0.6, 0.8), colours = pal_locuszoom("default")(5)) + 
    ylim(c(0,5)) + labs(y="-log10(pval)") +my_theme() +
    theme(axis.text.x=element_blank(), legend.position = "none")

chr2_GTEx$RS.ID <- factor(chr2_GTEx$RS.ID, levels=unique(chr2_GTEx$RS.ID))

p2 <- ggplot(chr2_GTEx, aes(x=RS.ID, y=-log10(P.value), fill=Tissue)) + 
    geom_col(position=position_dodge2()) + facet_wrap(vars(Gene.Symbol), ncol =2) +
    theme(axis.text.x =element_text(angle=90),
          axis.title.x=element_blank()) +
    my_theme()

p1  
p2  + labs(title="GTEx expression data for highlighted SNPs")
# cowplot::plot_grid(p1,p2, ncol=1, rel_heights = c(1,3), labels=c("A","B"))
```

### Chr14 76.1Mb Parent-of-origin folic acid interaction  

Many SNPs in this region had entries in GTEx v8. GTEx data is shown for the highlighted SNPs.

```{r, echo=FALSE}
load("../p_origin.Rdata")
load("../p_LD1.Rdata")

p_origin   <- p_origin[p_origin$region == "chr14\nr14",]
chr14_GTEx <- read.table("chr14_GTEx.txt", header=TRUE, sep='\t', quote="")
positions <- sapply(strsplit(chr14_GTEx[,3], ":"), "[[", 2)
positions <- as.numeric(positions)

chr14_GTEx <- chr14_GTEx[positions %in% p_origin$position,]
p_LD1 <- p_LD1[rownames(p_LD1) %in% p_origin$snp, colnames(p_LD1) %in% p_origin$snp]
lead  <- p_origin %>% slice_min(folate) %>% pull(snp)
p_origin$LD<- unname(p_LD1[,lead[1]]^2)

keep <- (-log10(p_origin$folate) > 2) & (p_origin$LD > 0.5)

p1 <- ggplot(p_origin, aes(position, -log10(folate))) +
  geom_point(color="grey") +
  geom_point(data=p_origin[keep,], color="red") +
  ylim(c(0,5)) + labs(y="-log10(pval)") +my_theme() +
  theme(axis.text.x=element_blank())

keep <- positions %in% p_origin$position[keep]
chr14_GTEx$RS.ID <- factor(chr14_GTEx$RS.ID, levels=unique(chr14_GTEx$RS.ID))

p2 <- ggplot(chr14_GTEx[keep,], aes(x=RS.ID, y=-log10(P.value), fill=Tissue)) + 
    geom_col(position=position_dodge2()) + facet_wrap(vars(Gene.Symbol)) +
    my_theme() + 
    theme(axis.text.x =element_text(angle=90),
          axis.title.x=element_blank(),
          legend.box.margin = margin(0,10,0,10)) +
   guides(fill = guide_legend(ncol = 3))

# legend <- cowplot::get_legend(p2)
# p2 <- cowplot::plot_grid(p2 + my_theme() + theme(legend.position = "none"), legend, ncol=1)

p1  
p2  + labs(title="GTEx expression data for highlighted SNPs")
# cowplot::plot_grid(p1,p2, ncol=1, rel_heights = c(1,3), labels=c("A","B"))
```





